---
title: "timepoint analysis"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(pacman)
p_load(tidyverse, ggsignif,multcomp, broom, emmeans, DescTools)
setwd("~/Documents/ESYT1_spinal_locomotion_analysis/Ventral video analysis/data") 
df_49<- read.table("speedProfile_p49.csv", sep="\t", header=T)
df_56<- read.table("speedProfile_p56.csv", sep="\t", header=T)
df_63<- read.table("speedProfile_p63.csv", sep="\t", header=T)
df_70<- read.table("speedProfile_p70.csv", sep="\t", header=T)
df_77<- read.table("speedProfile_p77.csv", sep="\t", header=T)
df_84<- read.table("speedProfile_p84.csv", sep="\t", header=T)
df_91<- read.table("speedProfile_p91.csv", sep="\t", header=T) 
df_98<- read.table("speedProfile_p98.csv", sep="\t", header=T)
df_105<- read.table("speedProfile_p105.csv", sep="\t", header=T)
df_112<- read.table("speedProfile_p112.csv", sep="\t", header=T)
df_112_237<- read.table("speedProfile_p112_237.csv", sep="\t", header=T)
df_49$time_point <- 49
df_56$time_point <- 56
df_63$time_point <- 63
df_70$time_point <- 70
df_77$time_point <- 77
df_84$time_point <- 84
df_91$time_point <- 91
df_98$time_point <- 98
df_105$time_point <- 105
df_112$time_point <- 112
df_112_237$time_point <- 112 #mouse #237 analyzed after rest of p112

#correcting cadence and stride, as video clip was slowed down for p112 #237
#Cadence
df_112_237$hLCad. <- df_112_237$hLCad.*5
df_112_237$hRCad. <- df_112_237$hRCad.*5
df_112_237$fLCad. <- df_112_237$fLCad.*5
df_112_237$fRCad <- df_112_237$fRCad*5

#Stride
df_112_237$hLStride <- df_112_237$hLStride/5
df_112_237$hRStride <-df_112_237$hRStride/5
df_112_237$fLStride <- df_112_237$fLStride /5
df_112_237$fRStride <- df_112_237$fRStride/5

#Bind all timepoint dataframes together
df <- rbind(df_49,df_56, df_63, df_70, df_77, df_84, df_91, df_98, df_105, df_112, df_112_237)

```


Genotyping
```{r}
df2 <- df
df2$Genotype <- 0
#Convert Genotype to SOD1En1cre
df2$Genotype <- ifelse(grepl('^19_', df$Name), 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(grepl('^50_', df$Name), 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(grepl('^237_', df$Name), 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(grepl('^239_', df$Name), 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(grepl('^318_', df$Name), 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(grepl('^44_', df$Name), 'SOD1En1cre', df2$Genotype)

#Convert Genotype to En1cre
df2$Genotype <- ifelse(grepl('^9_', df$Name), 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(grepl('^19E_', df$Name), 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(grepl('^33_', df$Name), 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(grepl('^80_', df$Name), 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(grepl('^42_', df$Name), 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(grepl('^48_', df$Name), 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(grepl('^62_', df$Name), 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(grepl('^66_', df$Name), 'En1cre', df2$Genotype)

#Convert Genotype to SOD1
df2$Genotype <- ifelse(grepl('^17_', df$Name), 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(grepl('^21_', df$Name), 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(grepl('^52_', df$Name), 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(grepl('^56_', df$Name), 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(grepl('^233_', df$Name), 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(grepl('^235_', df$Name), 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(grepl('^320_', df$Name), 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(grepl('^1_', df$Name), 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(grepl('^7_', df$Name), 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(grepl('^40_', df$Name), 'SOD1', df2$Genotype)

#Convert Genotype to wt
df2$Genotype <- ifelse(grepl('^11_', df$Name), 'wt', df2$Genotype)
df2$Genotype <- ifelse(grepl('^13_', df$Name), 'wt', df2$Genotype)
df2$Genotype <- ifelse(grepl('^15_', df$Name), 'wt', df2$Genotype)
df2$Genotype <- ifelse(grepl('^54_', df$Name), 'wt', df2$Genotype)
df2$Genotype <- ifelse(grepl('^316_', df$Name), 'wt', df2$Genotype)
df2$Genotype <- ifelse(grepl('^3_', df$Name), 'wt', df2$Genotype)
df2$Genotype <- ifelse(grepl('^5_', df$Name), 'wt', df2$Genotype)
df2$Genotype <- ifelse(grepl('^46_', df$Name), 'wt', df2$Genotype)
```

Give each mouse own number
```{r}

df2$mouse_number <- 0

#SOD1En1cre
df2$mouse_number <- ifelse(grepl('^19_', df$Name), '19', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^50_', df$Name), '50', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^237_', df$Name), '237', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^239_', df$Name), '239', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^318_', df$Name), '318', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^44_', df$Name), '44', df2$mouse_number)

#En1re
df2$mouse_number <- ifelse(grepl('^9_', df$Name), '9', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^19E_', df$Name), '19E', df2$mouse_number) 
df2$mouse_number <- ifelse(grepl('^33_', df$Name), '33', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^80_', df$Name), '80', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^42_', df$Name), '42', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^48_', df$Name), '48', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^62_', df$Name), '62', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^66_', df$Name), '66', df2$mouse_number)

#SOD1
df2$mouse_number <- ifelse(grepl('^17_', df$Name), '17', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^21_', df$Name), '21', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^52_', df$Name), '52', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^56_', df$Name), '56', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^233_', df$Name), '233', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^235_', df$Name), '235', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^320_', df$Name), '320', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^1_', df$Name), '1', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^7_', df$Name), '7', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^40_', df$Name), '40', df2$mouse_number)

#wt
df2$mouse_number <- ifelse(grepl('^11_', df$Name), '11', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^13_', df$Name), '13', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^15_', df$Name), '15', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^54_', df$Name), '54', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^316_', df$Name), '316', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^3_', df$Name), '3', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^5_', df$Name), '5', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('^46_', df$Name), '46', df2$mouse_number)
```


Removal of mouse 1, 7 and 80 from dataset as they were injected at different timepoints than the other mice.
```{r}
df2$mouse_number<- as.numeric(df2$mouse_number) #Change mouse number from class character to class numeric
df2<-df2[!(df2$mouse_number==80 | df2$mouse_number==1 | df2$mouse_number == 7),] 

```

Generating mean cadence and stride for each observation
```{r}
#Calculating mean cadence
df2$MeanCadence <- rowMeans(df2[,c("fLCad.","fRCad", "hLCad.", "hRCad.")])

#Calculating mean stride
df2$MeanStride <- rowMeans(df2[,c("hLStride","hRStride", "fLStride", "fRStride")])
```

Ordering Genotype variables by degree of manipulation (wt --> SOD1En1cre)
The purpose of this is for the bar plots to show the genotypes on the x-axis in order of how much the mouse is manipulated, starting from complete controls to intervention
```{r}
df2$Genotype <- factor(df2$Genotype, levels = c("wt", "En1cre", "SOD1", "SOD1En1cre"))
```

Creating datasets based on timepoint

```{r}
p49 <- newdata <- subset(df2, time_point==49)
p56 <- newdata <- subset(df2, time_point==56)
p63 <- newdata <- subset(df2, time_point==63)
p70 <- newdata <- subset(df2, time_point==70)
p77 <- newdata <- subset(df2, time_point==77)
p84 <- newdata <- subset(df2, time_point==84)
p91 <- newdata <- subset(df2, time_point==91)
p98 <- newdata <- subset(df2, time_point==98)
p105 <- newdata <- subset(df2, time_point==105)
p112 <- newdata <- subset(df2, time_point==112)
```

#Section used for creating excel files over genotypes and indivdiual mice, including mean stats for speed, acceleration, cadence and stride
Subset p49 by Genotype
```{r}
SOD1_p49 <- newdata <- subset(p49, Genotype=='SOD1')
SOD1En1cre_p49 <- newdata <- subset(p49, Genotype=='SOD1En1cre')
En1cre_p49 <- newdata <- subset(p49, Genotype=='En1cre')
wt_p49 <- newdata <- subset(p49, Genotype=='wt')
```

Subset p56 by Genotype
```{r}
SOD1_p56 <- newdata <- subset(p56, Genotype=='SOD1')
SOD1En1cre_p56 <- newdata <- subset(p56, Genotype=='SOD1En1cre')
En1cre_p56 <- newdata <- subset(p56, Genotype=='En1cre')
wt_p56 <- newdata <- subset(p56, Genotype=='wt')
```

Subset p63 by Genotype
```{r}
SOD1_p63 <- newdata <- subset(p63, Genotype=='SOD1')
SOD1En1cre_p63 <- newdata <- subset(p63, Genotype=='SOD1En1cre')
En1cre_p63 <- newdata <- subset(p63, Genotype=='En1cre')
wt_p63 <- newdata <- subset(p63, Genotype=='wt')
```

Subset p70 by Genotype
```{r}
SOD1_p70 <- newdata <- subset(p70, Genotype=='SOD1')
SOD1En1cre_p70 <- newdata <- subset(p70, Genotype=='SOD1En1cre')
En1cre_p70 <- newdata <- subset(p70, Genotype=='En1cre')
wt_p70 <- newdata <- subset(p70, Genotype=='wt')
```

Subset p77 by Genotype
```{r}
SOD1_p77 <- newdata <- subset(p77, Genotype=='SOD1')
SOD1En1cre_p77 <- newdata <- subset(p77, Genotype=='SOD1En1cre')
En1cre_p77 <- newdata <- subset(p77, Genotype=='En1cre')
wt_p77 <- newdata <- subset(p77, Genotype=='wt')
```

Subset p84 by Genotype
```{r}
SOD1_p84 <- newdata <- subset(p84, Genotype=='SOD1')
SOD1En1cre_p84 <- newdata <- subset(p84, Genotype=='SOD1En1cre')
En1cre_p84 <- newdata <- subset(p84, Genotype=='En1cre')
wt_p84 <- newdata <- subset(p84, Genotype=='wt')
```

Subset p91 by Genotype
```{r}
SOD1_p91 <- newdata <- subset(p91, Genotype=='SOD1')
SOD1En1cre_p91 <- newdata <- subset(p91, Genotype=='SOD1En1cre')
En1cre_p91 <- newdata <- subset(p91, Genotype=='En1cre')
wt_p91 <- newdata <- subset(p91, Genotype=='wt')
```

Subset p98 by Genotype
```{r}
SOD1_p98 <- newdata <- subset(p98, Genotype=='SOD1')
SOD1En1cre_p98 <- newdata <- subset(p98, Genotype=='SOD1En1cre')
En1cre_p98 <- newdata <- subset(p98, Genotype=='En1cre')
wt_p98 <- newdata <- subset(p98, Genotype=='wt')
```

Subset p105 by Genotype
```{r}
SOD1_p105 <- newdata <- subset(p105, Genotype=='SOD1')
SOD1En1cre_p105 <- newdata <- subset(p105, Genotype=='SOD1En1cre')
En1cre_p105 <- newdata <- subset(p105, Genotype=='En1cre')
wt_p105 <- newdata <- subset(p105, Genotype=='wt')
```

Subset p112 by Genotype
```{r}
SOD1_p112 <- newdata <- subset(p112, Genotype=='SOD1')
SOD1En1cre_p112 <- newdata <- subset(p112, Genotype=='SOD1En1cre')
En1cre_p112 <- newdata <- subset(p112, Genotype=='En1cre')
wt_p112 <- newdata <- subset(p112, Genotype=='wt')
```

Mean for each mouse

P49
```{r}
p49_means_mouse_by_Genotype <- p49 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)


```

P56
```{r}
p56_means_mouse_by_Genotype <- p56 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

P63
```{r}
p63_means_mouse_by_Genotype <- p63 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

P70
```{r}
p70_means_mouse_by_Genotype <- p70 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

p77
```{r}
p77_means_mouse_by_Genotype <- p77 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

p84
```{r}
p84_means_mouse_by_Genotype <- p84 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

p91
```{r}
p91_means_mouse_by_Genotype <- p91 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

p98
```{r}
p98_means_mouse_by_Genotype <- p98 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

p105
```{r}
p105_means_mouse_by_Genotype <- p105 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

p112
```{r}
p112_means_mouse_by_Genotype <- p112 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(PeakAcc. = mean(PeakAcc.), Avg.Speed = mean(Avg.Speed), MeanCadence = mean(MeanCadence), MeanStride = mean(MeanStride)) %>% 
  arrange(Genotype)

```

Dataframe with mean per mouse all timepoints

```{r}
df_mean_per_mouse <- rbind(p49_means_mouse_by_Genotype, p56_means_mouse_by_Genotype, p63_means_mouse_by_Genotype, p70_means_mouse_by_Genotype, p77_means_mouse_by_Genotype, p84_means_mouse_by_Genotype, p91_means_mouse_by_Genotype, p98_means_mouse_by_Genotype, p105_means_mouse_by_Genotype, p112_means_mouse_by_Genotype)
```


Create dataset with data from onset of locomotor phenotype

for mean
```{r}
onset_17 <-df_mean_per_mouse %>% filter(mouse_number =='17' & time_point == 70)
onset_19 <- df_mean_per_mouse%>% subset((mouse_number =='19' & time_point == 49))
onset_21 <- df_mean_per_mouse%>% subset((mouse_number =='21' & time_point == 63))
onset_50 <- df_mean_per_mouse%>% subset((mouse_number =='50' & time_point == 112))
onset_52 <- df_mean_per_mouse%>% subset((mouse_number =='52' & time_point == 77))
onset_56 <- df_mean_per_mouse%>% subset((mouse_number =='56' & time_point == 49))
onset_40 <- df_mean_per_mouse%>% subset((mouse_number =='40' & time_point == 91))
onset_44 <- df_mean_per_mouse%>% subset((mouse_number =='44' & time_point == 98))
onset_233 <- df_mean_per_mouse%>% subset((mouse_number =='233' & time_point == 56))
onset_235 <- df_mean_per_mouse%>% subset((mouse_number =='235' & time_point == 84))
onset_237 <- df_mean_per_mouse%>% subset((mouse_number =='237' & time_point == 112))
onset_239 <- df_mean_per_mouse%>% subset((mouse_number =='239' & time_point == 112))
onset_318 <- df_mean_per_mouse%>% subset((mouse_number =='318' & time_point == 70))
onset_320 <- df_mean_per_mouse%>% subset((mouse_number =='320' & time_point == 70))
onset_wt <- df_mean_per_mouse%>% subset((Genotype == 'wt' & time_point == 63))
onset_En1cre <- df_mean_per_mouse%>% subset((Genotype == 'En1cre' & time_point == 63))

onset_mean <- rbind(onset_17, onset_19, onset_21, onset_50, onset_52, onset_56, onset_40, onset_44, onset_233, onset_235, onset_237, onset_239, onset_318, onset_320, onset_wt, onset_En1cre)
```

for video triplicates
```{r}
onset_17 <-df2 %>% filter(mouse_number =='17' & time_point == 70)
onset_19 <- df2%>% subset((mouse_number =='19' & time_point == 49))
onset_21 <- df2%>% subset((mouse_number =='21' & time_point == 63))
onset_50 <- df2%>% subset((mouse_number =='50' & time_point == 112))
onset_52 <- df2%>% subset((mouse_number =='52' & time_point == 77))
onset_56 <- df2%>% subset((mouse_number =='56' & time_point == 49))
onset_40 <- df2%>% subset((mouse_number =='40' & time_point == 91))
onset_44 <- df2%>% subset((mouse_number =='44' & time_point == 98))
onset_233 <- df2%>% subset((mouse_number =='233' & time_point == 56))
onset_235 <- df2%>% subset((mouse_number =='235' & time_point == 84))
onset_237 <- df2%>% subset((mouse_number =='237' & time_point == 112))
onset_239 <- df2%>% subset((mouse_number =='239' & time_point == 112))
onset_318 <- df2%>% subset((mouse_number =='318' & time_point == 70))
onset_320 <- df2%>% subset((mouse_number =='320' & time_point == 70))
onset_wt <- df2%>% subset((Genotype == 'wt' & time_point == 63))
onset_En1cre <- df2%>% subset((Genotype == 'En1cre' & time_point == 63))

onset_all <- rbind(onset_17, onset_19, onset_21, onset_50, onset_52, onset_56, onset_40, onset_44, onset_233, onset_235, onset_237, onset_239, onset_318, onset_320, onset_wt, onset_En1cre)
```

#Plots

Set color palette for plots

For the plots, colors from the palette 'GrandBudapest1' from the 'wesanderson' package (Karthik Ram and Hadley Wickham (2018). wesanderson: A Wes Anderson Palette Generator. R package
  version 0.3.6. https://CRAN.R-project.org/package=wesanderson) were used, but colors re-ordered so the darkest color was applied to the wt mice
```{r}
color_palette = c("#5B1A18","#D67236","#FD6467", "#F1BB7B")
```

#Genotype-wise analyses

##Plots used in paper (onset and p112)

Preparing the ggsignif function to use Dunnet's p-values
```{r}

stars <- function(p) {
  case_when(
    p <= 0.001 ~ "***",
    p <= 0.01  ~ "**",
    p <= 0.05  ~ "*",
    TRUE       ~ "ns"
  )
}

```

Bar plots P112 only

```{r}

#speed

#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
p112$Genotype <- factor(p112$Genotype, levels = c("SOD1", "wt", "En1cre", "SOD1En1cre"))
anova_speed_1way_p112 <- aov(Avg.Speed ~ Genotype, data = p112)

Dunnet<- glht(anova_speed_1way_p112, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
p112$Genotype <- factor(p112$Genotype, levels = c("wt", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= p112, aes(x=Genotype, y=Avg.Speed)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
          geom_jitter(data=p112_means_mouse_by_Genotype, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.3)+
  ylim(0, 29) + ylab("Average speed cm/s")+
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
        ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(22,24,26)
  )



#Acceleration
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
p112$Genotype <- factor(p112$Genotype, levels = c("SOD1", "wt", "En1cre", "SOD1En1cre"))
anova_acceleration_1way_p112 <- aov(PeakAcc. ~ Genotype, data = p112)

Dunnet<- glht(anova_acceleration_1way_p112, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
p112$Genotype <- factor(p112$Genotype, levels = c("wt", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= p112, aes(x=Genotype, y=PeakAcc.)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
          geom_jitter(data=p112_means_mouse_by_Genotype, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.3)+
  ylim(0, 2.5) + ylab("Peak acceleration (cm/s^2)")+
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
           ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(2,2.2, 2.4)
  )


#Cadence
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
p112$Genotype <- factor(p112$Genotype, levels = c("SOD1", "wt", "En1cre", "SOD1En1cre"))
anova_cadence_1way_p112 <- aov(MeanCadence ~ Genotype, data = p112)

Dunnet<- glht(anova_cadence_1way_p112, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
p112$Genotype <- factor(p112$Genotype, levels = c("wt", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= p112, aes(x=Genotype, y=MeanCadence)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
      geom_jitter(data=p112_means_mouse_by_Genotype, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.3)+
  ylim(0, 5)+ ylab("Step frequency (Hz)")+
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
        ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(4.1,4.45,4.8)
  )

#Stride
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
p112$Genotype <- factor(p112$Genotype, levels = c("SOD1", "wt", "En1cre", "SOD1En1cre"))
anova_stride_1way_p112 <- aov(MeanStride ~ Genotype, data = p112)

Dunnet<- glht(anova_stride_1way_p112, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
p112$Genotype <- factor(p112$Genotype, levels = c("wt", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= p112, aes(x=Genotype, y=MeanStride)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
      geom_jitter(data=p112_means_mouse_by_Genotype, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.3)+
  ylim(0, 9) + ylab("Stride length (cm)")+
  theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(7,7.8,8.6)
  )

```
onset only

```{r}

#speed
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("SOD1", "wt", "En1cre", "SOD1En1cre"))
anova_speed_1way_onset <- aov(Avg.Speed ~ Genotype, data = onset_all)
summary(anova_speed_1way_onset)

Dunnet<- glht(anova_speed_1way_onset, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("wt", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= onset_all, aes(x=Genotype, y=Avg.Speed)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
      geom_jitter(data=onset_mean, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.3)+
  ylim(0, 32) + ylab("Average speed cm/s") +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
    ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(23,26,29)
  )

#Acceleration
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("SOD1", "wt", "En1cre", "SOD1En1cre"))
anova_acceleration_1way_onset <- aov(PeakAcc. ~ Genotype, data = onset_all)

Dunnet<- glht(anova_acceleration_1way_onset, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("wt", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= onset_all, aes(x=Genotype, y=PeakAcc.)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
          geom_jitter(data=onset_mean, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.3)+
  ylim(0, 3.3)+ ylab("Peak acceleration (cm/s^2)") +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
          ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(2.65, 2.9,3.15)
  )

#Cadence
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("SOD1", "wt", "En1cre", "SOD1En1cre"))
anova_cadence_1way_onset <- aov(MeanCadence ~ Genotype, data = onset_all)

Dunnet<- glht(anova_cadence_1way_onset, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("wt", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= onset_all, aes(x=Genotype, y=MeanCadence)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
          geom_jitter(data=onset_mean, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.3)+
  ylim(0, 6)+ ylab("Step frequency (Hz)") +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
        ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(4.7,5.1,5.5)
  )

#Stride
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("SOD1", "wt", "En1cre", "SOD1En1cre"))
anova_stride_1way_onset <- aov(MeanStride ~ Genotype, data = onset_all)


Dunnet<- glht(anova_stride_1way_onset, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("wt", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= onset_all, aes(x=Genotype, y=MeanStride)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
        geom_jitter(data=onset_mean, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.3)+
  ylim(0,9) + ylab("Stride length (cm)") +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(6,7,8)
  )


```

Longitudinal plot Speed
```{r}

ggplot(df2, aes(x=time_point, y= Avg.Speed, color=Genotype)) +
  geom_point(stat = "summary",fun.data=mean_se) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5) +
    scale_color_manual(values = color_palette)+ 
  geom_line(stat = "summary", fun.data=mean_se)+
    theme(panel.background = element_rect(fill = "white"), text = element_text(size=15), axis.line = element_line(color="black"), legend.key = element_rect(colour = "transparent", fill = "white")) +
  xlab("Postnatal day (timepoint)") + ylab("Average speed (cm/s)")

```

Peak acceleration
```{r}
ggplot(df2, aes(x=time_point, y= PeakAcc., color=Genotype)) +
  geom_point(stat = "summary",fun.data=mean_se) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5) +
    scale_color_manual(values = color_palette)+ 
  geom_line(stat = "summary",fun.data=mean_se)+
    theme(panel.background = element_rect(fill = "white"), text = element_text(size=15), axis.line = element_line(color="black"), legend.key = element_rect(colour = "transparent", fill = "white")) +
  xlab("Postnatal day (timepoint)") + ylab("Peak acceleration (cm/s^2)")

```

MeanCadence
```{r}
ggplot(df2, aes(x=time_point, y= MeanCadence, color=Genotype)) +
  geom_point(stat = "summary", fun.data=mean_se) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5) +
    scale_color_manual(values = color_palette)+ 
  geom_line(stat = "summary", fun.data=mean_se)+
    theme(panel.background = element_rect(fill = "white"), text = element_text(size=15), axis.line = element_line(color="black"), legend.key = element_rect(colour = "transparent", fill = "white")) +
  xlab("Postnatal day (timepoint)") + ylab("Step frequency (Hz)")

```

MeanStride
```{r}
ggplot(df2, aes(x=time_point, y= MeanStride, color=Genotype)) +
  geom_point(stat = "summary",fun.data=mean_se) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5) +
    scale_color_manual(values = color_palette)+ 
  geom_line(stat = "summary",fun.data=mean_se)+
    theme(panel.background = element_rect(fill = "white"), text = element_text(size=15), axis.line = element_line(color="black"), legend.key = element_rect(colour = "transparent", fill = "white")) +
  xlab("Postnatal day (timepoint)") + ylab("Stride length (cm)")

```


##Statistics

Two-way anova across timepoints
```{r}
#Ordering Genotype variables by manipulation so SOD1 becomes comparison group
df2$Genotype <- factor(df2$Genotype, levels = c("SOD1", "wt", "En1cre", "SOD1En1cre"))

df2$time_point <- as.factor(df2$time_point) #independent variables have to be recoded as factors to allow 'stratification' by timepoint for two-way anova

### Speed
anova_speed_2way <- aov(Avg.Speed ~ Genotype*time_point, data = df2)
summary(anova_speed_2way)

emmeans(anova_speed_2way,specs = trt.vs.ctrl~ Genotype|time_point)


### Cadence
anova_cadence_2way <- aov(MeanCadence ~ Genotype*time_point, data = df2)
summary(anova_cadence_2way)

emmeans(anova_cadence_2way,specs = trt.vs.ctrl~ Genotype|time_point)

### Stride
anova_stride_2way <- aov(MeanStride ~ Genotype*time_point, data = df2)
summary(anova_stride_2way)

emmeans(anova_stride_2way,specs = trt.vs.ctrl~ Genotype|time_point)


###Acceleration

anova_acceleration_2way <- aov(PeakAcc. ~ Genotype*time_point, data = df2)
summary(anova_acceleration_2way)

emmeans(anova_acceleration_2way,specs = trt.vs.ctrl~ Genotype|time_point)

```

one-way anova at p112

```{r}


#Ordering Genotype variables by manipulation so SOD1 becomes 'control' in Dunnet post-hoc tests
p112$Genotype <- factor(p112$Genotype, levels = c("SOD1", "wt", "En1cre", "SOD1En1cre"))


#speed
anova_speed_1way_p112 <- aov(Avg.Speed ~ Genotype, data = p112)
summary(anova_speed_1way_p112)
DunnettTest(x=p112$Avg.Speed, g=p112$Genotype)

#cadence
anova_cadence_1way_p112 <- aov(MeanCadence ~ Genotype, data = p112)
summary(anova_cadence_1way_p112)
DunnettTest(x=p112$MeanCadence, g=p112$Genotype)

#stride
anova_stride_1way_p112 <- aov(MeanStride ~ Genotype, data = p112)
summary(anova_stride_1way_p112)
DunnettTest(x=p112$MeanStride, g=p112$Genotype)

#acceleration
anova_acceleration_1way_p112 <- aov(PeakAcc. ~ Genotype, data = p112)
summary(anova_acceleration_1way_p112)
DunnettTest(x=p112$PeakAcc., g=p112$Genotype) #Dunnet's test

```

one-way anova at onset

```{r}

#Ordering Genotype variables by manipulation so SOD1 becomes 'control' in Dunnet post-hoc tests
onset_all$Genotype <- factor(onset_all$Genotype, levels = c("SOD1", "wt", "En1cre", "SOD1En1cre"))

#speed
anova_speed_1way_onset <- aov(Avg.Speed ~ Genotype, data = onset_all)
summary(anova_speed_1way_onset)
DunnettTest(x=onset_all$Avg.Speed, g=onset_all$Genotype)

#cadence
anova_cadence_1way_onset <- aov(MeanCadence ~ Genotype, data = onset_all)
summary(anova_cadence_1way_onset)
DunnettTest(x=onset_all$MeanCadence, g=onset_all$Genotype)

#stride
anova_stride_1way_onset <- aov(MeanStride ~ Genotype, data = onset_all)
summary(anova_stride_1way_onset)
DunnettTest(x=onset_all$MeanStride, g=onset_all$Genotype)

#acceleration
anova_acceleration_1way_onset <- aov(PeakAcc. ~ Genotype, data = onset_all)
summary(anova_acceleration_1way_onset)
DunnettTest(x=onset_all$PeakAcc., g=onset_all$Genotype) #Dunnet's test

```

Write the whole dataset into an excel or csv file
```{r}
setwd("~/Documents/ESYT1_spinal_locomotion_analysis/Ventral video analysis/")

write.csv(df_mean_per_mouse,"mice_means.csv", row.names = FALSE)
write.csv(df2,"mice_triplets.csv", row.names = FALSE)

```

